Calibrating the pixel-level Kepler imaging data with a causal data-driven model \\

### 2015-02-24 cpm.tex
typo
typos

@@

\begin{abstract}
In general, astronomical observations are affected by several kinds of noise, each with its own causal source; there is photon noise, stochastic source variability, and residuals coming from imperfect calibration of the detector or telescope. In particular, the precision of NASA Kepler photometry for exoplanet science—the most precise photometric measurements of stars ever made—appears to be limited by unknown or untracked variations in spacecraft pointing and temperature, and unmodeled stellar variability. Here we present the Causal Pixel Model (CPM) for Kepler data, a data-driven model intended to capture variability but preserve transit signals. The CPM works at the pixel level (not the photometric measurement level); it can capture more fine-grained information about the variation of the spacecraft than is available in the pixel-summed aperture photometry. The basic idea is that CPM predicts each target pixel value from a large number of pixels of other stars sharing the instrument variabilities while not containing any information on possible transits at the target star. In addition, we use the target star’s future and past (auto-regression). By appropriately separating the data into training and test sets, we ensure that information about any transit will be perfectly isolated from the fitting of the model. The method has four hyper-parameters (the number of predictor stars, the auto-regressive window size, and two L2-regularization amplitudes for model components), which we set by cross-validation. We determine a generic set of hyper-parameters that works well on most of the stars and apply the method to a corresponding set of target stars. For these stars, we calculate the Combined Differential Photometric Precision (CDPP), which indicate the noise level of a certain time scale. We find that we can consistently outperform (for the purposes of exoplanet detection) the Kepler Pre-search Data Conditioning (PDC) method for exoplanet discovery.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\columnwidth]{compare_10187017}
\caption{\label{hyperparameter} Comparsion between default and optimized hyper-parameters. The three panels plot show the results from default, In the top panel, SAP flux is plotted in black, CPM prediction of default hyper parameters is in red line, \name prediction of optimized hyper-parameters is in green line. The middle panel shows the CPM flux of the optimized hyper-parameters, while the bottom is the CPM flux of the default hyper-parameters, the two vertical blue lines indicate the location of the transit signal,  and signal to noise ratio is calculated in both situation. The optimized hyper-parameter performs better than the Default one, but the results from both situation are close. }
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\columnwidth]{distortion_9822284_normal}
\caption{\label{distortion} CPM preserves transit signals. Fake signal with amplitude 1.0005 is inserted into the single pixel (1, 2) of star KIC 9822284. In the top panel, the black points shows the SAP flux of the pixel with the injected signal, the green line is the fitting of the SAP flux, and the red line is the CPM prediction. In the bottom panel, red and green points represent the CPM flux and fit flux, while the signal value is indicated by the horizontal line. CPM largely preserves the signal while the fitting flux overfit the signal.}
\end{figure}

@@

To give a view on how our method performs, \name\ is applied on several stars with 
known transit signals with different variability and magnitudes 
as shown in Fig.~\ref{fluxes}. From the plots, we can see that \name\ works quite well
on these stars,  producing the \name\ fluxes with low variation and preserving the transit
signals.
In addition, we compare 6 quiet stars (non-variable stars) with the 
PDC method The results illustrate that for these quiet stars our approach removes 
a major part of the variability present in the PDC light curves, while preserving the transit signals. 
To provide a quantitative comparison, we ran CPM on 1000 stars from the whole Kepler input catalog (500 chosen randomly from the whole list, and 500 random G-type sun-like stars), and estimate the Combined Differential Photometric Precision (CDPP) for CPM and PDC (before CDPP is estimated, the 100 most variable stars in the list are
removed, since PDC is not designed to remove stellar variability). CDPP is an estimate of the relative precision in a time window, indicating the noise level seen by a transit signal with a given duration. The duration is typically chosen to be 3, 6, or 12 hours. We use the 12-hours CDPP metric, since the transit duration of an earth-like planet is roughly 10 hours. Fig.~\ref{cdpp} presents our CDPP comparison of CPM and PDC, showing that our method (CPM) outperforms PDC.

@@

\begin{figure*}[htb]
\centering
\begin{subfigure}[htb]{0.33\columnwidth}
\includegraphics[width=\columnwidth]{pdc_cdpp12_r}
\caption{}
\end{subfigure}%
%\hfill
\begin{subfigure}[htb]{0.33\columnwidth}
\includegraphics[width=\columnwidth]{boxplot_r}
\caption{}
\end{subfigure}%
%\hfill
\begin{subfigure}[htb]{0.33\columnwidth}
\includegraphics[width=\columnwidth]{pdc_cdpp12_hist_r100}
\caption{}
\end{subfigure}
\caption{\label{cdpp} Comparison of the proposed method (CPM) to the Kepler Pre-search Data Conditioning (PDC) method in terms of Combined Differential Photometric Precision (CDPP) (see text). Plot (a) shows our performance (red) vs.\ the PDC performance in a scatter plot, as a function of star magnitude (note that larger magnitude means fainter stars, and smaller values of CDPP indicate a higher quality as measured by CDPP.  Plot (b) bins the same dataset and shows box plots within each bin, indicating median, top quartile and bottom quartile. The red box corresponds to CPM, while the black box refers to PDC. Plot (c), finally, shows a histogram of CDPP values. Note that the red histogram has more mass towards the left, i.e., smaller values of CDPP, indicating that our method overall outperforms PDC, the Kepler ``gold standard.'' 
}
\end{figure*}

@@

\section{Discussion}
We have presented a a simple yet highly effective method that works in pixel-level based on the causal structure of the \Kepler\ data to calibrate the \Kepler\ light curves, which is intended to search 
exoplanets. However, based on our assumption of the causal structure of the \Kepler 
data, if we turn off the auto-regressive components and keep the predictors from other
stars working  in the \name, we should be able to still remove the systematics while
preserve the intrinsic stellar  variability, since there is no reason that these independent
stars can be used to predict the stellar variability of the target star. However, in fact, we
can not preserve the stellar variability just by turning off the auto-regressive 
components. One possible reason can be that, since our model is so flexible 
(thousands of parameters), in principle it can fit any functions. In this sense, we should 
run cross-validation with the auto-regressive components turned off and to find out how
many pixels from other stars we need to just remove systematics while preserve stellar
variability, but in fact we do not have a good criterion to check how well we preserve
stellar variability in cross-validation because we do not have enough knowledge of the 
true stellar variability. Thus for the time being, we just focus on producing well-
calibrated light curves for searching exoplanets, but in long term, we are looking
forward to find a method to preserve the stellar variability.

However, forget all the unsolved problems, we expect that our method will enable
astronomical discoveries at higher sensitivity on the existing \Kepler data as well as on
future missions.  As we know, \Kepler K2 data are suffering from the broken-down of
reaction wheels, which makes the satellite hard to control and introduce significant
systematics. Moreover, in 2017, NASA is planning the launch of another space 
telescope---TESS (Transiting Exoplanet Survey Satellite), which will perform an all-sky 
survey for small (earth-like) planets of nearby M stars. We think our method can be simply
extended for these project to help achieving higher precession and more scientific results.

>>>>

\usepackage{natbib}

@@

\begin{abstract}
In general, astronomical observations are affected by several kinds of noise, each with its own causal source; 
there is photon noise, stochastic source variability, and residuals coming from imperfect calibration of the detector or telescope. 
In particular, the precision of NASA Kepler photometry for exoplanet science—
the most precise photometric measurements of stars ever made—
appears to be limited by unknown or untracked variations in spacecraft pointing and temperature, 
and unmodeled stellar variability. Here we present the Causal Pixel Model (CPM) for Kepler data, 
a data-driven model intended to capture variability but preserve transit signals. 
The CPM works at the pixel level (not the photometric measurement level); 
it can capture more fine-grained information about the variation of the spacecraft than is available in the pixel-summed aperture photometry. 
The basic idea is that CPM predicts each target pixel value from a large number of pixels of other stars sharing the instrument variabilities 
while not containing any information on possible transits at the target star. In addition, we use the target star’s future and past (auto-regression). 
By appropriately separating the data into training and test sets, we ensure that information about any transit will be perfectly isolated from the fitting of the model. 
The method has four hyper-parameters (the number of predictor stars, the auto-regressive window size, 
and two L2-regularization amplitudes for model components), which we set by cross-validation. 
We determine a generic set of hyper-parameters that works well on most of the stars and apply the method to a corresponding set of target stars. 
For these stars, we calculate the Combined Differential Photometric Precision (CDPP), which indicates the noise level of a certain time scale. 
We find that we can consistently outperform (for the purposes of exoplanet detection) the Kepler Pre-search Data Conditioning (PDC) method for exoplanet discovery.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\columnwidth]{compare_10187017}
\caption{\label{hyperparameter} Comparsion between default and optimized hyper-parameters. 
In the top panel, SAP flux is plotted in black, \name\ prediction of default hyper parameters is in red line, 
\name\ prediction of optimized hyper-parameters is in green line. 
The middle panel shows the CPM flux of the optimized hyper-parameters, 
while CPM flux of the default hyper-parameters is in the bottom panel, 
the two vertical blue lines indicate the location of the transit signal, and signal to noise ratio is calculated in both situation. 
The optimized hyper-parameter performs better than the default one, but the results from both situation are close. }
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\columnwidth]{distortion_9822284_normal}
\caption{\label{distortion} \name\ preserves transit signals. 
Fake signal with amplitude 1.0005 is inserted into the single pixel (1, 2) of star KIC 9822284. 
In the top panel, the black points shows the SAP flux of the pixel with the injected signal, 
the green line is the fitting of the SAP flux, and the red line is the \name\ prediction. 
In the bottom panel, red and green points represent the \name\ flux and fit flux, 
while the signal value is indicated by the horizontal line. 
CPM largely preserves the signal while the fitting flux overfit the signal.}
\end{figure}

@@

To give a view on how our method performs, \name\ is applied on several stars with 
known transit signals but different variability and magnitudes as shown in Fig.~\ref{fluxes}. 
From the plots, we can see that \name\ works quite well on these stars, producing the \name\ fluxes 
with low variation and preserving the transit signals.
In addition, \name\ and PDC are compared based on 6 quiet stars 
(non-variable stars, first two rows of Fig.~\ref{fluxes}). 
The results illustrate that for these quiet stars our approach removes 
a major part of the variability present in the PDC light curves, while preserving the transit signals. 
To provide a quantitative comparison, we ran \name\ on 1000 stars from the whole Kepler input catalog 
(500 chosen randomly from the whole list, and 500 random G-type sun-like stars), 
and estimate the Combined Differential Photometric Precision (CDPP) for \name\ and PDC. 
CDPP is an estimate of the relative precision in a time window, 
indicating the noise level seen by a transit signal with a given duration. 
The duration is typically chosen to be 3, 6, or 12 hours. 
We use the 12-hours CDPP metric, since the transit duration of an earth-like planet is roughly 10 hours. 
Before CDPP is estimated, the 100 most variable stars in the list are removed, 
since PDC is not designed to remove stellar variability, and we want to compare \name\ and PDC on the same footing.
Fig.~\ref{cdpp} presents our CDPP comparison of \name\ and PDC, showing that our method (\name) outperforms PDC.

@@

\begin{figure*}[htb]
\centering
\begin{subfigure}[htb]{0.33\columnwidth}
\includegraphics[width=\columnwidth]{pdc_cdpp12_r}
\caption{}
\end{subfigure}%
%\hfill
\begin{subfigure}[htb]{0.33\columnwidth}
\includegraphics[width=\columnwidth]{boxplot_r}
\caption{}
\end{subfigure}%
%\hfill
\begin{subfigure}[htb]{0.33\columnwidth}
\includegraphics[width=\columnwidth]{pdc_cdpp12_hist_r100}
\caption{}
\end{subfigure}
\caption{\label{cdpp} Comparison of the proposed method (CPM) to the \Kepler\ Pre-search Data Conditioning (PDC) method 
in terms of Combined Differential Photometric Precision (CDPP) (see text). 
Plot (a) shows our performance (red) vs.\ the PDC performance in a scatter plot, as a function of star magnitude 
(note that larger magnitude means fainter stars, and smaller values of CDPP indicate a higher quality as measured by CDPP.  
Plot (b) bins the same dataset and shows box plots within each bin, indicating median, top quartile and bottom quartile. 
The red box corresponds to CPM, while the black box refers to PDC. Plot (c), finally, shows a histogram of CDPP values. 
Note that the red histogram has more mass towards the left, i.e., smaller values of CDPP, indicating that our method overall outperforms PDC, the Kepler ``gold standard.'' 
}
\end{figure*}

@@

\section{Discussion}
We have presented a simple yet highly effective method that works in pixel-level 
based on the causal structure of the \Kepler\ data to calibrate the \Kepler\ light curves, 
which is intended to search exoplanets. 
However, based on our assumption of the causal structure of the \Kepler\ 
data, if we turn off the auto-regressive components and keep the predictors from other
stars working  in the \name, we should be able to still remove the systematics while
preserve the intrinsic stellar  variability, since there is no reason that these independent
stars can be used to predict the stellar variability of the target star. However, in fact, we
can not preserve the stellar variability just by turning off the auto-regressive 
components. One possible reason can be that, since our model is so flexible 
(thousands of parameters), in principle it can fit any functions. In this sense, we should 
run cross-validation with the auto-regressive components turned off and to find out how
many pixels from other stars we need to just remove systematics while preserve stellar
variability, but in fact we do not have a good criterion to check how well we preserve
stellar variability in cross-validation because we do not have enough knowledge of the 
true stellar variability. Thus for the time being, we just focus on producing well-
calibrated light curves for searching exoplanets, but in long term, we are looking
forward to find a method to preserve the stellar variability.

However, forget all the unsolved problems, we expect that our method will enable
astronomical discoveries at higher sensitivity on the existing \Kepler\ data as well as on
future missions.  As we know, \Kepler\ K2 data are suffering from the broken-down of
reaction wheels, which makes the satellite hard to control and introduce significant
systematics. Moreover, in 2017, NASA is planning the launch of another space 
telescope---TESS (Transiting Exoplanet Survey Satellite), which will perform an all-sky 
survey for small (earth-like) planets of nearby M stars. We think our method can be simply
extended for these project to help achieving higher precession and more scientific results.
